---
title: "R package coga: The Description of Algorithm and Correctness"
author: "Chaoran Hu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Description of Algorithm and Correctness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(coga)
if (!requireNamespace("microbenchmark", quietly = TRUE)) {
  stop("Pkg microbenchmark needed for this vignette to work. Please install it.",
    call. = FALSE)
}
```

# Introduction

This vignette is the description of algorithm in this package and the correctness check via simulation experiment. The algorithm of this package comes from Moschopoulos Peter G. (1985). This vignette also give some useful informations.

# Algorithm

Assume that we have several random variables, $X_1, ..., X_n$, and all random variables follow gamma distribution independently with shape parameters $\alpha_i$ and scale parameters $\beta_i$, where $i = 1, ..., n$. Then, the density of $Y = X_1 + ... + X_n$ can be expressed as:

$$g(y) = C \sum_{k=0}^{\infty} \lambda_k y^{\rho + k - 1} e^{-y/\beta_1} / (\Gamma(\rho + k) \beta_{1}^{\rho + k})$$


And the distribution function $G(w)=Pr(Y<w)$ is expressed as:

$$G(w) = C \sum_{k=0}^{\infty} \lambda_k \int_{0}^{w} (y^{\rho + k - 1} e^{-y/\beta_1} / (\Gamma(\rho + k) \beta_{1}^{\rho + k})) dy$$

The integrate in this formula is incomplete gamma function and can be calculated by distribution function of gamma distribution.

More details about this algorithm can be found in paper of Moschopoulos Peter G. (1985).

# Correctness

Assume that we have two random variables, $X_1$ and $X_2$, where $X_1$ is a gamma distribution with shape parameter $3$, and rate parameter $2$, and $X_2$ is a gamma distribution with shape parameter $4$, and rate parameter $3$. The density and distribution funciton of $Y = X_1 + X_2$ will be calculated.

Correctness check for density function:

```{r, echo=TRUE, fig.width = 7, fig.height = 6}
y <- rcoga(1000000, c(3,4), c(2,3))
grid <- seq(0, 8, length.out=1000)
pdf <- dcoga(grid, shape=c(3, 4), rate=c(2, 3))
 
plot(density(y), col="blue")
lines(grid, pdf, col="red")
```

Correctness check for distribution function:

```{r, echo=TRUE, fig.width = 7, fig.height = 6}
y <- rcoga(1000000, c(3,4), c(2,3))
grid <- seq(0, 8, length.out=1000)
cdf <- pcoga(grid, shape=c(3, 4), rate=c(2, 3))

plot(ecdf(y), col="blue")
lines(grid, cdf, col="red")
```

# Speed

The 'dcoga' and 'pcoga' functions in this package 'coga' is based on Cpp code. The following experiment shows the advantage of Cpp code, which runs on a Intel(R) Core(TM) i7-4870HQ CPU @ 2.50GHz computer.

```{r, echo=TRUE}
grid <- seq(0, 15, length.out=10)

microbenchmark::microbenchmark(
    dcoga(grid, shape=c(3,4,5), rate=c(2,3,4)),
    coga:::dcoga.R(grid, shape=c(3,4,5), rate=c(2,3,4)),
    pcoga(grid, shape=c(3,4,5), rate=c(2,3,4)),
    coga:::pcoga.R(grid, shape=c(3,4,5), rate=c(2,3,4))
)
```

Note: In this example, 'dcoga.R', and 'pcoga.R' are the R version functions for density, and distribution functions of convolution of gamma distributions. We do not put these two R functions as export functions in package 'coga', but you can still use them by 'coga:::dcoga', and 'coga:::pcoga'.

# Parameters Recycling

Please take care of that `dcoga`, and `pcoga` in this package can handle different lengths of parameter `shape` and `rate` by recycling shorter parameter. That means that `dcoga(3, c(2,3), c(3,4,5,3,4))` and `dcoga(3, c(2,3,2,3,2), c(3,4,5,3,4))` will give the same result.

# References

[1] Moschopoulos, Peter G. "The distribution of the sum of independent gamma random variables." Annals of the Institute of Statistical Mathematics 37.1 (1985): 541-544.
